Are Transformed Low Discrepancy Points Also Low Discrepancy?

### 2019-06-02 InvPts_verB.tex
Updates from Overleaf

\FJH{Fred does not understand this example.}
A simple example of integral approximation below shows that a naively inverse transformed sequence $\Udes$ could result in very misleading estimate even if $D(\Udes,F_\unif, K)$ is small.

\emph{Example}: Consider estimating the integral
\begin{align}\label{eq:badexample}
\int_{\mathbb{R}^d} \frac{x_1+\cdots+x_d}{1+10^{-8}(x_1^2+\cdots+x_d^2)}\dif \Phi(\vx) & = \Ex\left[\frac{\sum\limits_{i=1}^d x_i}{1+10^{-8}\|\vx\|_2^2}\right],\,\,\,\vx\sim N({\bf 0}, {\mathsf I}_d)
\end{align}
using Monte Carlo method, with $d=10$ and $N=2^{10}$ points. Analytically, the integral \eqref{eq:badexample} equals 0 since $\frac{x_1+\cdots+x_d}{1+10^{-8}(x_1^2+\cdots+x_d^2)}$ is an odd function in each dimension and the probability density function of standard normal is an even function. We start with two uniform random sequences. One is a Sobol sequence that has small uniform discrepancy, denoted as $\Udes_1$. Then we move the closest point to origin $0^d$ in $\Udes_1$ to $\left(10^{-15}\right)^d$, and denote the new sequence as $\Udes_2$. We compare the integral estimate with samples $\vPsi(\Udes_1)$ and $\vPsi(\Udes_2)$. Table \ref{tab:badexample} shows that the uniform discrepancies of $\Udes_1$ and $\Udes_2$ don't differ much, but the normal discrepancy of $\vPsi(\Udes_2)$ is far larger than that of $\vPsi(\Udes_1)$. In return, the estimate error corresponding to $\Udes_2$ is about 60 times larger than that corresponding to $\Udes_1$.

% Table generated by Excel2LaTeX from sheet 'Sheet1'
\begin{table}[htbp]
  \centering
  \caption{Comparison of Integral Estimate}
    \begin{tabular}{ccccc}
    \hline
   $\Udes$       & $D(\Udes,F_\unif, K)$     & $D(\vPsi(\Udes),F_\normal, K)$     & Estimate \\
    \hline
    $\Udes_1$     &   0.0285    &  18.57     & 0.0011 \\
    $\Udes_2$     &  0.0292      &  58.82     & -0.0660 \\
    \hline
    \end{tabular}%
  \label{tab:badexample}%
\end{table}%

>>>>

\FJH{Changing example.}

\begin{example} A simple cubature example illustrates that a naively inverse transformed low discrepancy design, $\Udes$, may yield a large $D(\vPsi(\Udes),F_\normal, K)$ and also a large cubature error. Consider the integration problem in \eqref{eq:integration} with
\begin{subequations} \label{eq:badexample}
\begin{gather}
\vX \sim  N(\vzero, {\mathsf I}_d), \qquad f(\vx) = \frac{x_1^2+\cdots+x^2_d}{1+10^{-8}(x_1^2+\cdots+x_d^2)}, \qquad Y = f(\vX),\\
\mu = \Ex(Y) = \int_{\reals^d} \frac{x_1^2+\cdots+x^2_d}{1+10^{-8}(x_1^2+\cdots+x_d^2)} \phi(\vx) \, \dif \vx
\end{gather}
\end{subequations}
where $\phi$ is the probability density function for the standard multivariate normal distribution.
Since the integrand above is a function of $x_1^2 + \cdots + x_d^2$, $\mu$ can be written as a one dimensional integral.  For $d=10$, $\mu = 10$ to at least $15$ significant digits using quadrature.

We can also approximate the integral in \eqref{eq:badexample} using a $d=10$, $N=512$ cubature \eqref{eq:cubature}.  We compare cubatures using two designs.  The design $\Xdes_1$ is the inverse normal transformation of a scrambled Sobol' sequence, $\Udes_1$, which has low discrepancy with respect to the uniform distribution on the $d$-dimensional unit cube.  The design $\Udes_2$ takes the point in $\Udes_1$ that is closet to $\vzero$ and moves it to $\left(10^{-15}, \ldots, 10^{-15}\right)$, which is very close to $\vzero$.  As seen in Table \ref{tab:badexample}, the two uniform designs have quite similar, small discrepancies.  However, the transformed designs, $\Xdes_j = \vPsi(\Udes_j)$ for $j=1,2$, have much different discrepancies with respect to the normal distribution.  This is due to the point in $\Xdes_2$ that has large negative coordinates.  Furthermore, the cubatures, $\hat{\mu}$, based on these two designs have significantly different errors.  The first design has both a smaller discrepancy and a smaller error than the second.  This could not have been induced by looking at the discrepancies of the original uniform designs.

% Table generated by Excel2LaTeX from sheet 'Sheet1'
\begin{table}[htbp]
  \centering
  \caption{Comparison of Integral Estimate}
    \begin{tabular}{ccccc}
   $\Udes$       & $D(\Udes,F_\unif, K)$     & $D(\vPsi(\Udes),F_\normal, K)$     & $\hat{\mu}$  & Relative Error\\
    \toprule
    $\Udes_1$     &   0.0285    &  18.57     & 10.0182 & 0.0018\\
    $\Udes_2$     &  0.0292      &  58.82     & 11.2238 & 0.1224
    \end{tabular}%
  \label{tab:badexample}%
\end{table}%

% % Table generated by Excel2LaTeX from sheet 'Sheet1'
% \begin{table}[htbp]
%   \centering
%   \caption{Comparison of Integral Estimate}
%     \begin{tabular}{ccccc}
%     \hline
%   $\Udes$       & $D(\Udes,F_\unif, K)$     & $D(\vPsi(\Udes),F_\normal, K)$     & Estimate \\
%     \hline
%     $\Udes_1$     &   0.0285    &  18.57     & 0.0011 \\
%     $\Udes_2$     &  0.0292      &  58.82     & -0.0660 \\
%     \hline
%     \end{tabular}%
%   \label{tab:badexample}%
% \end{table}%

\end{example}
